Bushfire Ignition in Victoria

### 2016-01-22 numbat_pres/numbatpres.Rmd

---
title: "Bushfires in R"
author: "Elle Saber"
date: "20 January 2016"
output: beamer_presentation
---

## Modelling Probability of Ignition of Bushfires in Victoria

Data Processing involved with the honours project

## Slide with Bullets
```{r mysize=TRUE, size='small'}
library(ncdf)
temp.nc <- open.ncdf("./TMAX1984-2009.nc")
temp.nc

@@

## Fire Map

![Map of Fire Occurence](figures/Data/firemap.pdf)

## NCDF Temp Data random Day

![Max Temperature](figures/Data/tmax.pdf)

## NCDF time series

![Time series Silvan](figures/Data/silvan.pdf)

>>>>

---
title: "Bushfire Ignition in Victoria"
author: "Elle Saber "
highlighter: highlight.js
output: pdf_document
job: null
knit: slidify::knit2slides
mode: selfcontained
hitheme: tomorrow
subtitle: An adventure in wrangling spatial data
framework: io2012
widgets: mathjax
---

```{r, fig.width=12,echo=FALSE, fig.align='center'}
library(png)
library(grid)
img <- readPNG("/Users/ellesaber/bushfire/numbat_pres/cfa.png")
 grid.raster(img)
```

---

# The Fire Danger Rating System

The McArthur Forest Fire Danger Index:

$$latex F=2.0 exp(-0.450 + 0.987 \ln(D)-0.0345H+0.0338T+0.0234V) $$

```{r, fig.width=7,echo=FALSE, fig.align='center'}
library(png)
library(grid)
img <- readPNG("/Users/ellesaber/Documents/HonsData/RMD/fdr-chart.png")
 grid.raster(img)
```

---

# WRF Model Data in NetCDF Format

```{r message=FALSE}
library(ncdf)
temp.nc <- open.ncdf("./tmax_subset.nc")
temp.nc
```

---

# WRF Data Time Series - Silvan

```{r, fig.width=12,echo=FALSE, fig.align='center'}
library(ggplot2)
library(tidyr)
load("/Users/ellesaber/Documents/HonsData/analysis/silvan.rda")

silv_long <- gather(silvan, key = var, value = val, col=c(4,8:11))
ggplot(silv_long, aes(x=date, y=val)) + geom_line(color="darkblue") + 
  facet_grid(var~., scales="free") + theme_bw() + 
  scale_color_brewer(palette = "Dark2") +
  ggtitle("Meteorological Variables") + 
  xlab("Date") + 
  ylab("Value")

@@

---

# WRF Model Temperature

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=12}
library(ncdf)
library(raster)
library(rasterVis)
sub.temp <-open.ncdf("./tmax_subset.nc")
ras <- raster("./tmax_subset.nc")
myTheme=rasterTheme(region=brewer.pal('Reds', n=9))

levelplot(ras, margin=FALSE, par.settings=myTheme)
```

---

# Shapefiles

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=12}

# layerName is the name of the unzipped shapefile without file type extensions 
layerName <- "STE11aAust"  
# Read in the data
aus.shape <- readOGR(dsn="/Users/ellesaber/bushfire/data_raw/shps", layer=layerName) 
#remove other polygons to leave only Victoria
vic.shp <- aus.shape[aus.shape$STATE_NAME == "Victoria",]
plot(vic.shp)

---

# ABS Statistical Divisions

```{r, echo = FALSE, message = FALSE, fig.align = 'center', warning = FALSE, fig.width=12}
# layerName is the name of the unzipped shapefile without file type extensions 
layerName <- "SD11aAust"  
# Read in the data
aus.shape <- readOGR(dsn="/Users/ellesaber/bushfire/data_raw/shps", layer=layerName) 
#remove other polygons to leave only Victoria
vic.shp <- aus.shape[aus.shape$STATE_CODE == 2,]
plot(vic.shp)
```

---

# Fire & Incident Reporting Service

```{r, fig.width=13,echo=FALSE, fig.align='center'}

library(png)
library(grid)
img <- readPNG("/Users/ellesaber/bushfire/numbat_pres/firs.png")
 grid.raster(img)
```

---

# Australian Map Grid Zones

```{r, fig.width=10,echo=FALSE, fig.align='center'}
img <- readPNG("/Users/ellesaber/bushfire/numbat_pres/zonemap.png")
 grid.raster(img)

---
# Fire Map

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=12}
library(ggplot2)
library(data.table)
library(ggmap)
library(sp)
library(rgdal)
library(dplyr)
library(tidyr)

#datset
load("/Users/ellesaber/Documents/HonsData/fire_occur.rda")
#mappy stuff

box <- make_bbox(long, lat, fire.dt, f=0.05)
myMap <- get_stamenmap(bbox=box, zoom=6, maptype = "toner", crop=TRUE)
#map of vic
vicmap <- ggmap(myMap,
      extent = "device")
#add in fires!
firemap <- vicmap +geom_point(aes(x=long, y=lat), data=fire.dt, alpha=.5, color="red3")
firemap

---&twocol

# Grassland Curing & Vapour Pressure 
*** =left
```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=6}
library(raster)
library(rasterVis)
library(RColorBrewer)
library(gridExtra)

gs <- raster("/Users/ellesaber/Documents/HonsData/grasslandcuring/file20090125-20090201.nc", varname="curing")

plot(gs, col=bpy.colors(10))

*** =right

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=6}
library(raster)
library(rasterVis)
library(RColorBrewer)
library(gridExtra)

gs <- raster("/Users/ellesaber/Documents/HonsData/grasslandcuring/file20090125-20090201.nc", varname="curing")

e <- extent(gs) #upper and lower limits to contain VIC

ps <- raster("/Users/ellesaber/Documents/HonsData/vapour/20090207.nc")
ps <- crop(ps, e)

plot(ps, cm.colors(10))

---

# Grassland Curing Evolution

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=13}
library(ggplot2)
library(dplyr)

load("/Users/ellesaber/bushfire/robject/dat0.rda")

K <- dat0 %>% 
  group_by(date,div) %>% 
  summarise( avg=mean(curing, na.rm=TRUE), max=max(curing, na.rm=TRUE), min=min(curing, na.rm=TRUE))
K <- gather(K, key=var, value = val, 3:5)
vec <- c("2000-10-01", "2001-10-01", "2002-10-01", "2003-10-01", "2003-10-01", "2004-10-01",
         "2005-10-01", "2006-10-01", "2007-10-01", "2008-10-01", "2009-10-01")
vec <- as.Date(vec)

K$seas <- findInterval(K$date, vec)
K$seas <- ordered(K$seas,
                  levels = c(1:11),
                  labels = c("2000", "2001", "2002", "2003", "2004", "2005", 
                             "2006", "2007", "2008", "2009", "2010"))

K <- filter(K, seas=="2008")

ggplot(K, aes(x=date, y=val, color=var))+
  geom_line()+facet_wrap(~div) +
  ggtitle("Grassland Curing over 2008 Fire Season") + 
  xlab("Date") + 
  ylab("Grassland Curing Value")

---

# The Fire Danger Index and Fire Occurence

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=13}
library(ggplot2)

load("/Users/ellesaber/bushfire/robject/sub.rda")
newsub$fire <- as.factor(newsub$fire)

ggplot(newsub, aes(ffdi, color=fire)) + geom_density() + facet_wrap(~div) + scale_x_log10() +   scale_color_brewer(palette = "Dark2")

---

# The Fire Danger Index and Temperature

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=13}

library(ggplot2)

load("/Users/ellesaber/bushfire/robject/sub.rda")
newsub$fire <- as.factor(newsub$fire)

ggplot(newsub, aes(ffdi, temp, color=fire)) + geom_density2d() + facet_wrap(~div) +   scale_color_brewer(palette = "Dark2")
```

---

# The Fire Danger Index and Grassland Curing

```{r, echo=F, message=F, fig.align='center', warning=FALSE, fig.width=13}
ggplot(newsub, aes(curing, ffdi, color=fire)) + geom_density2d() + facet_wrap(~div) +   scale_color_brewer(palette = "Dark2")

---
# Data sets

- WRF Model
- CFA Fire & Indcident Reporting Database
- CFA Grassland Curing
- BOM Vapour Pressure & Precipitation

# Key R Packages

- dplyr & tidyr 
- rvest
- NCDF
- raster
- sp
- rgdal
